### 2020-05-19 writeup/proof_transfer_ex.tex
revise

\section{Supplementary Materials for Section \ref{sec_insight}}

@@

We state several helper lemmas to get a bound on the variance of $\hat{\beta}_t^{\STL}$ and $\hat{\beta}_t^{\MTL}$.
The first lemma, which is a folklore result in random matrix theory, helps determine the asymptotic limit of $\te(\hat{\beta}_t^{\STL})$, as $p$ goes to infinity.
\begin{lemma}\label{lem_minv}[\todo{ref?}]
	Let $X\in\real^{n\times p}$ be a random matrix that contains i.i.d. row vectors with mean $0$ and covariance $\Sigma\in\real^{p\times p}$.
	In the setting when $n = c p$ we have that as $p$ goes to infinity,
	\[ \bigtr{(X^{\top}X)^{-1}\Sigma} = \frac{1}{c - 1}. \]
\end{lemma}

The second lemma, which deals the inverse of the sum of two random matrices, can be viewed as a special case of Theorem \ref{thm_model_shift}.
\begin{lemma}\label{lem_cov_shift}
	Let $X_i\in\real^{n_i\times p}$ be a random matrix that contains i.i.d. row vectors with mean $0$ and variance $\Sigma_i\in\real^{p\times p}$, for $i = 1, 2$.
	Denote by $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ and let $\lambda_1, \lambda_2, \dots, \lambda_p$ be the singular values of $M^{\top}M$ in decreasing order.
	When $n_1 = c_1 p$ and $n_2 = c_2 p$, we have that with high probability over the randomness of $X_1$ and $X_2$, the following equation holds
	\begin{align}%\label{lem_cov_shift_eq}
		\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2} = \frac{1}{c_1+c_2}\cdot \bigtr{ \frac{1}{p}\cdot(a_1 M^\top M + a_2)^{-1}} +\bigo{ n^{-1/2+\epsilon}},
	\end{align}
	for any constant $\epsilon>0$, where $a_1, a_2$ are solutions to the following deterministic equations:
	\begin{align}
		a_1 + a_2 = 1- \frac{1}{c_1 + c_2},~ a_1 + \frac{1}{p}\cdot\sum_{i=1}^p \frac{a_1}{(c_1 + c_2)(a_1 + a_2/ \lambda_i^2)} = \frac{c_1}{c_1 + c_2}. \label{eq_a12}
	\end{align}
\end{lemma}
We will give the proof of Lemma \ref{lem_cov_shift} in Section \ref{sec_maintools}.

@@

Now we prove the first statement of Proposition \ref{prop_data_size}. Notice that the function
$$ \frac{c_1(c_1+c_2)(c_2-1)}{(c_1+c_2-1)^2} =(c_2-1) \left(1 +\frac{c_2-2}{c_1}+\frac{1}{c_1(c_1+c_2)}\right)^{-1} $$
is strictly increasing with respect to $c_1$ as long as $c_2\ge 3$. In particular, for $c_1\ge 100$, we get the bound by taking $c_1\to \infty$:
$$\frac{c_1(c_1+c_2)(c_2-1)}{(c_1+c_2-1)^2} \cdot \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4 < 1.1^4 (c_2-1).$$
 Hence we conclude that for $c_1\ge 100$, we always have $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$ whp as long as $pd^2 \le \frac{2\sigma^2}{3(c_2-1)}$. This gives the first statement.

The second statement can be proved in a similar way. Suppose $c_1 \ge 110$ and $pd^2 > \frac{3\sigma^2}{2(c_2-1)}$. If $c_1 > \frac{(c_2-2) \sigma^2}{2(c_2 - 1) pd^2/3 - \sigma^2}$, we have
\begin{align*}
 pd^2 \cdot \frac{c_1(c_1+c_2)(c_2-1)}{(c_1+c_2-1)^2}  \cdot \bigbrace{1 - \sqrt{\frac{1}{c_1}}}^4 > \frac{2}{3} pd^2(c_2-1)\cdot \frac{c_1 }{c_1+(c_2-2)}  >\sigma^2, 
 \end{align*}
 where in the first step we used that 
 $$ \frac{(c_1+c_2)(c_1+c_2-2)}{(c_1+c_2-1)^2}  \cdot \bigbrace{1 - \sqrt{\frac{1}{c_1}}}^4 > \frac23$$
 for $c_1 \ge 110$. This shows that \eqref{neg1} holds as long as $p$ is large enough, and hence $\te(\hat{\beta}_t^{\MTL})> \te(\hat{\beta}_t^{\STL})$ holds. On the other hand, if $100\le c_1 \le \frac{(c_2-2) \sigma^2}{3(c_2 - 1) pd^2/2 - \sigma^2}$, then we have 
 \begin{align*}
 pd^2 \cdot \frac{c_1(c_1+c_2)(c_2-1)}{(c_1+c_2-1)^2}  \cdot \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4 < \frac{3}{2} pd^2(c_2-1)\cdot \frac{c_1 }{c_1+(c_2-2)}  < \sigma^2, 
 \end{align*}
 where in the first step we used that 
 $$ \frac{(c_1+c_2)(c_1+c_2-2)}{(c_1+c_2-1)^2}  \cdot \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4 < \frac32$$
 for $c_1 \ge 100$. This shows that \eqref{pos1} holds as long as $p$ is large enough, and hence $\te(\hat{\beta}_t^{\MTL})< \te(\hat{\beta}_t^{\STL})$ holds.
 \end{proof}

@@

	We have that the test error satisfies
\be\label{approxteM}  te(M)\left(1 -  \frac{n_2}{n_1-p} \frac{1}{\lambda_p^2 + \frac{n_2}{n_1-p}}\right)  \le  \frac{\sigma^2}{n_1+n_2}\tr\left( \frac{1}{a_1M^\top M + a_2}\right) \le te(M),\ee
where $\lambda_p$ is the smallest singular value of $p$ and
$$te(M):= \frac{\sigma^2}{a_1(n_1+n_2)}\tr\left( \frac{1}{M^\top M}\right) .$$
Moreover, for all $M$ satisfying \eqref{GMcons}, the minimum of $te(M)$ is attained when $M= a\id$.

From equation \eqref{eq_a2}, we get
$$a_1\ge \frac{n_1-p}{n_1+n_2},\quad a_2\le \frac{n_2}{n_1+n_2}.$$
With these two bounds, we can easily conclude \eqref{approxteM}.

>>>>

\section{Supplementary Materials for Section \ref{sec_insight}}\label{app_proof_main}

@@

We state several helper lemmas to get a bound on the variance of $\hat{\beta}_t^{\STL}$ and $\hat{\beta}_t^{\MTL}$.
The first lemma, which is a folklore result in random matrix theory, helps determine the asymptotic limit of $\te(\hat{\beta}_t^{\STL})$, as $p$ goes to infinity.

\begin{lemma}\label{lem_minv}[\todo{ref?}]
	Let $X\in\real^{n\times p}$ be a random matrix that contains i.i.d. row vectors with mean $0$ and covariance $\Sigma\in\real^{p\times p}$.
	In the setting when $n = c p$ we have that as $p$ goes to infinity,
	\[ \bigtr{(X^{\top}X)^{-1}\Sigma} = \frac{1}{c - 1}. \]
\end{lemma}

The second lemma, which deals the inverse of the sum of two random matrices, can be viewed as a special case of Theorem \ref{thm_model_shift}.

\begin{lemma}\label{lem_cov_shift}
	Let $X_i\in\real^{n_i\times p}$ be a random matrix that contains i.i.d. row vectors with mean $0$ and variance $\Sigma_i\in\real^{p\times p}$, for $i = 1, 2$.
	Denote by $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ and let $\lambda_1, \lambda_2, \dots, \lambda_p$ be the singular values of $M^{\top}M$ in decreasing order.
	When $n_1 = c_1 p$ and $n_2 = c_2 p$, we have that with high probability over the randomness of $X_1$ and $X_2$, the following equation holds
	\begin{align}%\label{lem_cov_shift_eq}
		\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2} = \frac{1}{c_1+c_2}\cdot \bigtr{ \frac{1}{p}\cdot(a_1 M^\top M + a_2)^{-1}} +\bigo{ n^{-1/2+\epsilon}},
	\end{align}
	for any constant $\epsilon>0$, where $a_1, a_2$ are solutions to the following deterministic equations:
	\begin{align}
		a_1 + a_2 = 1- \frac{1}{c_1 + c_2},~ a_1 + \frac{1}{p}\cdot\sum_{i=1}^p \frac{a_1}{(c_1 + c_2)(a_1 + a_2/ \lambda_i^2)} = \frac{c_1}{c_1 + c_2}. \label{eq_a12}
	\end{align}
\end{lemma}
We will give the proof of Lemma \ref{lem_cov_shift} in Section \ref{sec_maintools}.

@@

Now we prove the first statement of Proposition \ref{prop_data_size}. Notice that the function
$$ \frac{c_1(c_1+c_2)(c_2-1)}{(c_1+c_2-1)^2} =(c_2-1) \left(1 +\frac{c_2-2}{c_1}+\frac{1}{c_1(c_1+c_2)}\right)^{-1} $$
is strictly increasing with respect to $c_1$ as long as $c_2\ge 3$. In particular, by taking $c_1\to \infty$, we get the bound:
$$\frac{c_1(c_1+c_2)(c_2-1)}{(c_1+c_2-1)^2} \cdot \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4 <  (c_2-1) \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4.$$
 Hence we conclude that $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$ whp as long as 
 $$pd^2 +\oo(1) \le \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^{-4}\frac{\sigma^2}{c_2-1} .$$ 
 This gives the first statement by taking $c_1 > a$.

The second statement can be proved in a similar way. Suppose $c_1 > a$ and 
$pd^2 > (1 - a^{-1/2})^{-4}\frac{\sigma^2}{c_2-1}$. If $c_1 > \frac{(c_2-2) \sigma^2}{(1-a^{-1/2})^4(1-(a+c_2-2)^{-2})(c_2 - 1) pd^2 - \sigma^2}$, we have
\begin{align*}
 & pd^2 \cdot \frac{c_1(c_1+c_2)(c_2-1)}{(c_1+c_2-1)^2}  \cdot \bigbrace{1 - \sqrt{\frac{1}{c_1}}}^4  \\
 &>  \bigbrace{1 - \sqrt{\frac{1}{c_1}}}^4 \left( 1-\frac{1}{(c_1+c_2-2)^2}\right)\cdot \frac{ pd^2(c_2-1) c_1 }{c_1+(c_2-2)}  >\sigma^2\cdot (1+\oo(1)) , 
 \end{align*}
 where in the first step we used that 
 $$ \frac{(c_1+c_2)(c_1+c_2-2)}{(c_1+c_2-1)^2}   > 1-\frac{1}{(c_1+c_2-2)^2}.$$
This shows that \eqref{neg1} holds as long as $p$ is large enough, and hence $\te(\hat{\beta}_t^{\MTL})> \te(\hat{\beta}_t^{\STL})$ holds. On the other hand, if $c_1 < \frac{(c_2-2)\sigma^2}{(1+a^{-1/2})^4(c_2 - 1) pd^2 - \sigma^2}$, then we have 
 \begin{align*}
 pd^2 \cdot \frac{c_1(c_1+c_2)(c_2-1)}{(c_1+c_2-1)^2}  \cdot \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4 < \bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4 \cdot \frac{pd^2(c_2-1) c_1 }{c_1+(c_2-2)}  < \sigma^2\cdot (1-\oo(1)).
 \end{align*}
 This shows that \eqref{pos1} holds as long as $p$ is large enough, and hence $\te(\hat{\beta}_t^{\MTL})< \te(\hat{\beta}_t^{\STL})$ holds.
 \end{proof}

@@

......... We have that the test error satisfies
\be\label{approxteM}  te(M)\left(1 -  \frac{n_2}{n_1-p} \frac{1}{\lambda_p^2 + \frac{n_2}{n_1-p}}\right)  \le  \frac{\sigma^2}{n_1+n_2}\tr\left( \frac{1}{a_1M^\top M + a_2}\right) \le te(M),\ee
where $\lambda_p$ is the smallest singular value of $p$ and
$$te(M):= \frac{\sigma^2}{a_1(n_1+n_2)}\tr\left( \frac{1}{M^\top M}\right) .$$
Moreover, for all $M$ satisfying \eqref{GMcons}, the minimum of $te(M)$ is attained when $M= a\id$.

From equation \eqref{eq_a2}, we get
$$a_1\ge \frac{n_1-p}{n_1+n_2},\quad a_2\le \frac{n_2}{n_1+n_2}.$$
With these two bounds, we can easily conclude \eqref{approxteM}. Then solving equations \eqref{eq_a3} and \eqref{eq_a4}, we get that
\begin{align*}
a_3= \frac{ \frac{1}{n_2}\sum_{i=1}^p \frac{a_2^2 }{ (  a_2 + \lambda_i^2a_1)^2  }\left(1 - \frac{1}{n_1} \sum_{i=1}^p \frac{\lambda_i^4 a_1^2  }{  (a_2 + \lambda_i^2a_1)^2  }\right) +\left( \frac{1}{n_1}\sum_{i=1}^p \frac{\lambda_i^2 a_1^2}{  (a_2 + \lambda_i^2a_1)^2  }\right) \left(\frac{1}{n_2}\sum_{i=1}^p \frac{  \lambda_i^2 a_2^2 }{ (  a_2 + \lambda_i^2a_1)^2  }\right) } {\left(1- \frac1{n_2}\sum_{i=1}^p \frac{a_2^2}{ (a_2 + \lambda_i^2a_1)^2  }\right) \left(1 - \frac{1}{n_1} \sum_{i=1}^p \frac{\lambda_i^4 a_1^2  }{  (a_2 + \lambda_i^2a_1)^2  }\right) - \left(\frac{1}{n_2}\sum_{i=1}^p \frac{  \lambda_i^2 a_2^2 }{ (  a_2 + \lambda_i^2a_1)^2  }\right) \left(\frac{1}{n_1}\sum_{i=1}^p \frac{\lambda_i^2 a_1^2 }{  (a_2 + \lambda_i^2a_1)^2  }\right)} ,
\end{align*}
\begin{align*}
a_4= \frac{\frac{1}{n_1}\sum_{i=1}^p \frac{\lambda_i^2 a_1^2}{  (a_2 + \lambda_i^2a_1)^2  }}{\left(1- \frac1{n_2}\sum_{i=1}^p \frac{a_2^2}{ (a_2 + \lambda_i^2a_1)^2  }\right) \left(1 - \frac{1}{n_1} \sum_{i=1}^p \frac{\lambda_i^4 a_1^2  }{  (a_2 + \lambda_i^2a_1)^2  }\right) - \left(\frac{1}{n_2}\sum_{i=1}^p \frac{  \lambda_i^2 a_2^2 }{ (  a_2 + \lambda_i^2a_1)^2  }\right) \left(\frac{1}{n_1}\sum_{i=1}^p \frac{\lambda_i^2 a_1^2 }{  (a_2 + \lambda_i^2a_1)^2  }\right)} ,
\end{align*}

### 2020-05-19 writeup/transfer.tex
fig 2
revise

\begin{proposition}\label{prop_data_size}
	In the setting of Example \ref{ex_basic}, assume that $c_1 \ge 100$ and $c_2  \ge 3$ are fixed constants.
	We have the following conditions to determine whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$:
	\begin{itemize}
		\item If $d^2 < \frac{2\sigma^2}{3p (c_2 - 1)}$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$, for any $c_1$ larger than $??$.
		\item If $d^2 > \frac{3\sigma^2}{2p (c_2 - 1)}$, then we have the following transition depending on $c_1$:
		\begin{itemize}
			\item If $100 \le c_1 < \frac{(c_2-2)\sigma^2}{3(c_2 - 1) pd^2/2 - \sigma^2}$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
			\item If $c_1 > \frac{c_2 \sigma^2}{2(c_2 - 1)p d^2/3 - \sigma^2}$ and $c_1\ge 110$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
		\end{itemize}
	\end{itemize}
\end{proposition}

@@

\begin{proposition}\label{prop_covariate}
	In the setting of Example \ref{ex_basic}, assume that $d^2 < \sigma^2 / (c_1 p)$ and $c_1 > ??$.
	As a parameter of $M \in \cS_{\mu}$, we have that $(\te(\hat{\beta}_t^{\MTL}))(M)$ as a function of $M$ satisfies that 
$$ (\te(\hat{\beta}_t^{\MTL}))(\mu \id) \le \inf_{M\in \cal S_{\mu}}(\te(\hat{\beta}_t^{\MTL}))(M) + {\cor\text{some errors}}.$$	
	%is minimized when $M = \mu\id$.
\end{proposition}

>>>>

\begin{proposition}\label{prop_data_size}
	In the setting of Example \ref{ex_basic}, assume that $c_2  \ge 3$ is fixed and $c_1 > a$ for some fixed integer $a$.	We have the following conditions to determine whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$:
	\begin{itemize}
\item If $d^2 \le (1 + a^{-1/2})^{-4}\frac{\sigma^2}{p(c_2-1)}$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.

\item If $d^2 > (1 - a^{-1/2})^{-4}\frac{\sigma^2}{p (c_2 - 1)}$, then we have the following transition depending on $c_1$:
		\begin{itemize}
			\item If $c_1 < \frac{(c_2-2)\sigma^2}{(1+a^{-1/2})^4(c_2 - 1) pd^2 - \sigma^2}$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
			\item If $c_1 > \frac{(c_2-2) \sigma^2}{(1-a^{-1/2})^4(1-(a+c_2-2)^{-2})(c_2 - 1) pd^2 - \sigma^2}$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
		\end{itemize}
	\end{itemize}
\end{proposition}

@@

\begin{proposition}\label{prop_covariate}
	In the setting of Example \ref{ex_basic}, assume that $d^2 < \sigma^2 / (c_1 p)$ and $c_1 > ??$.
	As a parameter of $M \in \cS_{\mu}$, we have that $(\te(\hat{\beta}_t^{\MTL}))(M)$ as a function of $M$ satisfies that 
$$ (\te(\hat{\beta}_t^{\MTL}))(\mu \id) \le \left[1+ {\cor\text{some errors}}\right]\cdot\inf_{M\in \cal S_{\mu}}(\te(\hat{\beta}_t^{\MTL}))(M) .$$	
	%is minimized when $M = \mu\id$.
\end{proposition}
